---
title: "IMSE548_HW1_EMG"
author: "Shiva Rasouli"
date: "2024-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```
# Load libraries
```{r}
library(signal)
library(readr)
library(tidyverse)
```

# Read data
```{r}
data1 <- read_csv("../Hand sEMG dataset/1_raw.csv")


# Renaming columns for each dataset
colnames(data1) <- c("ch1", "ch2", "ch3", "ch4")

```


#EMG Plots for Participant1
```{r}
#sampling frequency
fs<-2000
time <- seq(from = 0, by = 1/fs, length.out = nrow(data1))
data1$time<-time

# Load the necessary libraries
library(tidyverse)

# Convert data from wide to long format
data_long <- data1 %>%
  pivot_longer(
    cols = starts_with("ch"),  
    names_to = "Channel",  
    values_to = "Signal"  
  )
data_long$Channel <- factor(data_long$Channel,
                            labels = c("Channel 1", "Channel 2", "Channel 3", "Channel 4"))
# Plot the data with ggplot2
raw_data <- ggplot(data_long, aes(x = time, y = Signal)) +
  geom_line(color="skyblue") + 
  facet_wrap(~Channel, scales = "free_y", ncol = 1) +
  theme_minimal() + # Use a minimal theme
  labs(title = " ", # Overall plot title
  x = "Time (s)", y = "EMG (mV)") + 
  theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), # Add grid lines
        panel.grid.minor = element_line(color = "white")) +
  geom_hline(yintercept = 0, linetype = "dashed") 

# Print the plot
print(raw_data)



ggsave("../plot/raw_plot.png", plot = raw_data, dpi = 300, width = 8, height = 6)

```

# FFT- to find cutoff frequencies
# NOTE: this section takes a long time to run
```{r}
# Calculate FFT
fft_data <- fft(data1[, 1:4] <- sapply(data1[, 1:4], as.numeric)) # Applying FFT to the first four columns excluding time

# Calculate Magnitude Spectrum
mag_spectrum <- Mod(fft_data)

# Frequency Vector
L <- nrow(data1) # Data length
fs <- 2000 # Sampling frequency
f <- seq(0, fs/2, length.out = length(mag_spectrum)/2) # Frequency vector for the first half


```
# Plot FFT
```{r}
# Plot without default y-axis, adjusting axis title sizes with cex.lab
plot(f, mag_spectrum[1:length(f)]/1e+06, type = 'l',  # Scale down the magnitude by 1e+06
     xlab = "Frequency (Hz)",
     ylab = expression(Magnitude ~ (10^6)),  # Use expression to format ylab
     main = "", col = "black", yaxt = "n",
     cex.lab = 0.8)  # Decrease the size of axis titles

# Add a custom y-axis with adjusted labels
axis(side = 2, at = seq(0, max(mag_spectrum)/1e+06, by = 2),  # Adjust 'by' as needed
     labels = seq(0, max(mag_spectrum)/1e+06, by = 2))

# Enhancing the plot with grid
grid()


```


# Band-pass filter
```{r}
#Lowe and upper end cutoff
lower_cutoff<-10
upper_cutoff<-500

#Filte order
order<-4  

# Normalize frequencies by Nyquist frequency
Wn <- c(lower_cutoff, upper_cutoff) / (fs / 2) 

# Design band-pass filter
filter_coefs <- butter(order, Wn, type = 'pass') 

#Apply the filter

filtered_data <- apply(data1, 2, function(channel_data) {
  filtfilt(filter_coefs, channel_data)
})



```

#Ploting data after Band Pass filtering 
```{r}
# Convert 'filtered_data' to a data frame
filtered_df <- as.data.frame(filtered_data)

filtered_df$time <- data1$time
filtered_df <- filtered_df %>%
  pivot_longer(
    cols = starts_with("ch"),  
    names_to = "Channel",  
    values_to = "Signal"  
  )
filtered_df$Channel <- factor(filtered_df$Channel,
                            labels = c("Channel 1", "Channel 2", "Channel 3", "Channel 4"))
# Plot the data with ggplot2
bp <- ggplot(filtered_df, aes(x = time, y = Signal)) +
  geom_line(color="skyblue") + 
  facet_wrap(~Channel, scales = "free_y", ncol = 1) +
  theme_minimal() + 
  labs(title = " ", 
  x = "Time (s)", y = "EMG (mV)") + 
  theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10),
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white")) +
  geom_hline(yintercept = 0, linetype = "dashed") 

# Print the plot
print(bp)


ggsave("../plot/filtered_plot.png", plot = bp, dpi = 300, width = 8, height = 6)
```


# Signal Processing: Full-wave rectification
```{r}
rectified_signal <- abs(filtered_data)
```

# Ploting signal after band-pass filtering and rectification
```{r}
library(ggplot2)
library(tidyr) # For pivot_longer()
# Convert MVC_Normalized to a dataframe
rectified_signal <- as.data.frame(rectified_signal)

# Convert the dataset to a long format
rectified_signal$time <- time

# Ensure tidyverse (or at least tidyr) is loaded
library(tidyverse)

# Pivot to long format
rectified_long <- rectified_signal %>% 
  pivot_longer(
    cols = -time,  
    names_to = "Channel",  
    values_to = "Signal"  
  )
# Update Channel to be a factor with levels like "Channel 1", "Channel 2", etc.
rectified_long$Channel <- factor(rectified_long$Channel,
                            labels = c("Channel 1", "Channel 2", "Channel 3", "Channel 4"))

# Plotting with faceting and custom channel labels
p <- ggplot(rectified_long, aes(x = time, y = Signal)) +
  geom_line(color="skyblue") + 
  theme_minimal() + 
  labs(title = "", 
  x = "Time (s)", y = "EMG (mV)") + 
  facet_wrap(~ Channel, ncol = 1, scales = "free_y") + 
  theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white")) +
  geom_hline(yintercept = 0, linetype = "dashed") 

print(p)
ggsave("../plot/rectified_plot.png", plot = p, dpi = 300, width = 8, height = 6)


```
#RMS envelope
```{r}
RMS <- matrix(0, nrow(data1), 4)  
window <- 100  # Window size

# Load necessary library
library(zoo)

for (i in 1:4) {
   RMS[, i] <- sqrt(rollapply(rectified_signal[, i]^2, window, mean, fill = NA, align = 'right'))
}

max_rms_values <- apply(RMS, MARGIN = 2, max, na.rm = TRUE)

print(round(max_rms_values,3))
```
#Plot RMS
```{r}
RMS <- as.data.frame(RMS)

RMS$time <- time


# Pivot to long format
RMS_long <- RMS %>% 
  pivot_longer(
    cols = -time,  
    names_to = "Channel",  
    values_to = "Signal"  
  )
# Update Channel to be a factor with levels like "Channel 1", "Channel 2", "Channel 3", "Channel 4"
RMS_long$Channel <- factor(RMS_long$Channel,
                            labels = c("Channel 1", "Channel 2", "Channel 3", "Channel 4"))

# Plotting with faceting and custom channel labels
p <- ggplot(RMS_long, aes(x = time, y = Signal)) +
  geom_line(color="skyblue") + # Add lines
  theme_minimal() + # Use a minimal theme
  labs(title = "RMS EMG over Time", # Overall plot title
  x = "Time (s)", y = "EMG RMS (mV)") + # Add labels
  facet_wrap(~ Channel, ncol = 1, scales = "free_y") + 
  theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white")) +
  geom_hline(yintercept = 0, linetype = "dashed") 

print(p)
ggsave("../plot/RMS_plot.png", plot = p, dpi = 300, width = 8, height = 6)

```



# Normalization (comparing with MVC)

```{r}
MVC <- c(1.757 ,0.931, 1.501 ,0.168)

MVC_Normalized <- matrix(NA, nrow = nrow(RMS), ncol = 4)

for (i in 1:4) {
  MVC_Normalized[, i] <- (RMS[, i] / MVC[i])
}

# Convert MVC_Normalized to a dataframe
MVC_Normalized <- as.data.frame(MVC_Normalized)

# Now add the 'time' column to the dataframe
MVC_Normalized$time <- time


# Pivot to long format
MVC_Normalized_long <- MVC_Normalized %>% 
  pivot_longer(
    cols = -time,  
    names_to = "Channel",  
    values_to = "Signal"  
  )
# Update Channel to be a factor with levels like "Channel 1", "Channel 2", etc.
MVC_Normalized_long$Channel <- factor(MVC_Normalized_long$Channel,
                            labels = c("Channel 1", "Channel 2", "Channel 3", "Channel 4"))

```

#Plotting normalized envelope against rectified signal

```{r}
library(dplyr)

# Convert the Time column to numeric
rectified_long$time <- as.numeric(as.character(rectified_long$time))
MVC_Normalized_long$time <- as.numeric(as.character(MVC_Normalized_long$time))
 
# Ensure that the Channel names are consistent between the two data frames
rectified_long$Channel <- as.character(rectified_long$Channel)
MVC_Normalized_long$Channel <- as.character(MVC_Normalized_long$Channel)
 
# Merge the data on Time and Muscle
combined_data <- full_join(rectified_long, MVC_Normalized_long, by = c("time", "Channel"))
combined_data <- combined_data %>%
  filter(time >= 0 & time <= 104)
# Now plot the data with 4 subplots for the non-percent data only
final_plot <- ggplot(combined_data, aes(x = time)) +
  geom_line(aes(y = Signal.x, colour = "R")) +
  geom_line(aes(y = Signal.y, colour = "N"), size=0.01) +
  facet_wrap(~Channel, scales = "free_y", ncol = 1) +
  theme_minimal()+
  scale_colour_manual(values = c("R" = "skyblue", "N" = "red"), name = NULL) +
  labs(title = "",
       x = "Time",
       y = "EMG (mV)") +
  
  theme_minimal() +
   theme(plot.title = element_text(hjust = 0.5), # Center the plot title
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white")) +
  geom_hline(yintercept = 0, linetype = "dashed") 

# Print the final plot
print(final_plot)
ggsave("../plot/two_plot.png", plot = final_plot, dpi = 300, width = 8, height = 6)

```

#Statistcal analysis: Compute the mean and median frequency
```{r}



colnames(MVC_Normalized) <- c("Channel1", "Channel2", "Channel3", "Channel4")
colnames(MVC_Normalized)[5] <- "time"

library(dplyr)

# Function to calculate mean and median for every 2000 rows for a given channel
calculate_stats <- function(data, channel_name) {
  data %>%
    mutate(row_group = ceiling(row_number() / 2000)) %>%
    group_by(row_group) %>%
    summarise(mean = mean(!!sym(channel_name), na.rm = TRUE),
              median = median(!!sym(channel_name), na.rm = TRUE)) %>%
    select(-row_group)
}

# Apply the function to each channel
stats_channel1 <- calculate_stats(MVC_Normalized, "Channel1")
stats_channel2 <- calculate_stats(MVC_Normalized, "Channel2")
stats_channel3 <- calculate_stats(MVC_Normalized, "Channel3")
stats_channel4 <- calculate_stats(MVC_Normalized, "Channel4")

# Combine the results into a new dataframe
mnf_mdf <- data.frame(Channel1_Mean = stats_channel1$mean,
                      Channel1_Median = stats_channel1$median,
                      Channel2_Mean = stats_channel2$mean,
                      Channel2_Median = stats_channel2$median,
                      Channel3_Mean = stats_channel3$mean,
                      Channel3_Median = stats_channel3$median,
                      Channel4_Mean = stats_channel4$mean,
                      Channel4_Median = stats_channel4$median)


# Calculate the total number of entries required for the 'time' sequence
total_segments <- nrow(mnf_mdf)

# Create a time sequence from 1 to 640, repeating it enough times to cover all segments
time_sequence <- rep(1:640, length.out = total_segments)

# Add the time sequence to the mnf_mdf dataframe
mnf_mdf$time <- time_sequence



```

#Plot mnf, and mdf for channel 1 for extention gesture
```{r}
library(ggplot2)

# Filter the first 60 rows for plotting
mnf_mdf_subset <- mnf_mdf[11:24, ]

# Create the plot
mnf<-ggplot(mnf_mdf_subset, aes(x = time)) + 
  geom_line(aes(y = Channel1_Mean, color = "Mean")) + 
  geom_line(aes(y = Channel1_Median, color = "Median")) +
  scale_color_manual(values = c("Mean" = "blue", "Median" = "red"), name=NULL) +
  labs(title = "",  #Channel 1 Mean and Median over Time for Participant1
       x = "Time (s)",
       y = "Value",
       color = "Legend") +
  theme_minimal()+
theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white"))


ggsave("../plot/mnf_1_1.png", plot = mnf, dpi = 300, width = 8, height = 6)


```
#Plot mnf, and mdf for channel 3 for extention gesture
```{r}
library(ggplot2)

# Filter the first 60 rows for plotting
mnf_mdf_subset <- mnf_mdf[11:24, ]

# Create the plot
mnf<-ggplot(mnf_mdf_subset, aes(x = time)) + 
  geom_line(aes(y = Channel3_Mean, color = "Mean")) + 
  geom_line(aes(y = Channel3_Median, color = "Median")) +
  scale_color_manual(values = c("Mean" = "blue", "Median" = "red"), name=NULL) +
  labs(title = "", 
       x = "Time (s)",
       y = "Value",
       color = "Legend") +
  theme_minimal()+
theme(plot.title = element_text(hjust = 0.5), # Center the plot title
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), # Smaller x-axis title
        axis.title.y = element_text(size = 10), # Smaller y-axis title
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white"))


ggsave("../plot/mnf_3_1.png", plot = mnf, dpi = 300, width = 8, height = 6)

```
#Plot mnf, and mdf for channel 2 for Flexion gesture
```{r}
library(ggplot2)

# Filter the first 60 rows for plotting
mnf_mdf_subset <- mnf_mdf[21:34, ]

# Create the plot
mnf<-ggplot(mnf_mdf_subset, aes(x = time)) + 
  geom_line(aes(y = Channel2_Mean, color = "Mean")) + 
  geom_line(aes(y = Channel2_Median, color = "Median")) +
  scale_color_manual(values = c("Mean" = "blue", "Median" = "red"), name=NULL) +
  labs(title = "",  
       x = "Time (s)",
       y = "Value",
       color = "Legend") +
  theme_minimal()+
theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white"))


ggsave("../plot/mnf_2_2.png", plot = mnf, dpi = 300, width = 8, height = 6)


```
#Plot mnf, and mdf for channel 4 for Flexion gesture
```{r}
library(ggplot2)

# Filter the first 60 rows for plotting
mnf_mdf_subset <- mnf_mdf[21:34, ]

# Create the plot
mnf<-ggplot(mnf_mdf_subset, aes(x = time)) + 
  geom_line(aes(y = Channel4_Mean, color = "Mean")) + 
  geom_line(aes(y = Channel4_Median, color = "Median")) +
  scale_color_manual(values = c("Mean" = "blue", "Median" = "red"), name=NULL) +
  labs(title = "",  
       x = "Time (s)",
       y = "Value",
       color = "Legend") +
  theme_minimal()+
theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white"))


ggsave("../plot/mnf_4_2.png", plot = mnf, dpi = 300, width = 8, height = 6)

```
#Plot mnf, and mdf for channel 2 for Ulnar Deviation gesture
```{r}
library(ggplot2)

# Filter the first 60 rows for plotting
mnf_mdf_subset <- mnf_mdf[31:44, ]

# Create the plot
mnf<-ggplot(mnf_mdf_subset, aes(x = time)) + 
  geom_line(aes(y = Channel2_Mean, color = "Mean")) + 
  geom_line(aes(y = Channel2_Median, color = "Median")) +
  scale_color_manual(values = c("Mean" = "blue", "Median" = "red"), name=NULL) +
  labs(title = "",  
       x = "Time (s)",
       y = "Value",
       color = "Legend") +
  theme_minimal()+
theme(plot.title = element_text(hjust = 0.5), # Center the plot title
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"),
        panel.grid.minor = element_line(color = "white"))


ggsave("../plot/mnf_2_3.png", plot = mnf, dpi = 300, width = 8, height = 6)


```
#Plot mnf, and mdf for channel 1for Ulnar Deviation gesture
```{r}
library(ggplot2)

# Filter the first 60 rows for plotting
mnf_mdf_subset <- mnf_mdf[31:44, ]

# Create the plot
mnf<-ggplot(mnf_mdf_subset, aes(x = time)) + 
  geom_line(aes(y = Channel1_Mean, color = "Mean")) + 
  geom_line(aes(y = Channel1_Median, color = "Median")) +
  scale_color_manual(values = c("Mean" = "blue", "Median" = "red"), name=NULL) +
  labs(title = "",  
       x = "Time (s)",
       y = "Value",
       color = "Legend") +
  theme_minimal()+
theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1),
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white"))


ggsave("../plot/mnf_1_3.png", plot = mnf, dpi = 300, width = 8, height = 6)

```
#Plot mnf, and mdf for channel 3 for Radias Deviation gesture
```{r}
library(ggplot2)

# Filter the first 60 rows for plotting
mnf_mdf_subset <- mnf_mdf[41:54, ]

# Create the plot
mnf<-ggplot(mnf_mdf_subset, aes(x = time)) + 
  geom_line(aes(y = Channel3_Mean, color = "Mean")) + 
  geom_line(aes(y = Channel3_Median, color = "Median")) +
  scale_color_manual(values = c("Mean" = "blue", "Median" = "red"), name=NULL) +
  labs(title = "",  
       x = "Time (s)",
       y = "Value",
       color = "Legend") +
  theme_minimal()+
theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white"))


ggsave("../plot/mnf_3_4.png", plot = mnf, dpi = 300, width = 8, height = 6)


```
#Plot mnf, and mdf for channel 1 for Radias Deviation gesture
```{r}
library(ggplot2)

# Filter the first 60 rows for plotting
mnf_mdf_subset <- mnf_mdf[41:54, ]

# Create the plot
mnf<-ggplot(mnf_mdf_subset, aes(x = time)) + 
  geom_line(aes(y = Channel1_Mean, color = "Mean")) + 
  geom_line(aes(y = Channel1_Median, color = "Median")) +
  scale_color_manual(values = c("Mean" = "blue", "Median" = "red"), name=NULL) +
  labs(title = "",  
       x = "Time (s)",
       y = "Value",
       color = "Legend") +
  theme_minimal()+
theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white"))


ggsave("../plot/mnf_4_4.png", plot = mnf, dpi = 300, width = 8, height = 6)

```
#Plot mnf, and mdf for channel 1 for Grip gesture
```{r}
library(ggplot2)

# Filter the first 60 rows for plotting
mnf_mdf_subset <- mnf_mdf[51:64, ]

# Create the plot
mnf<-ggplot(mnf_mdf_subset, aes(x = time)) + 
  geom_line(aes(y = Channel1_Mean, color = "Mean")) + 
  geom_line(aes(y = Channel1_Median, color = "Median")) +
  scale_color_manual(values = c("Mean" = "blue", "Median" = "red"), name=NULL) +
  labs(title = "",  
       x = "Time (s)",
       y = "Value",
       color = "Legend") +
  theme_minimal()+
theme(plot.title = element_text(hjust = 0.5), # Center the plot title
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white"))


ggsave("../plot/mnf_1_5.png", plot = mnf, dpi = 300, width = 8, height = 6)


```
#Plot mnf, and mdf for channel 2 for Grip gesture
```{r}
library(ggplot2)

# Filter the first 60 rows for plotting
mnf_mdf_subset <- mnf_mdf[51:64, ]

# Create the plot
mnf<-ggplot(mnf_mdf_subset, aes(x = time)) + 
  geom_line(aes(y = Channel2_Mean, color = "Mean")) + 
  geom_line(aes(y = Channel2_Median, color = "Median")) +
  scale_color_manual(values = c("Mean" = "blue", "Median" = "red"), name=NULL) +
  labs(title = "",  
       x = "Time (s)",
       y = "Value",
       color = "Legend") +
  theme_minimal()+
theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10),
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white"))


ggsave("../plot/mnf_2_5.png", plot = mnf, dpi = 300, width = 8, height = 6)

```
#Plot mnf, and mdf for channel 3 for Grip gesture
```{r}
library(ggplot2)

# Filter the first 60 rows for plotting
mnf_mdf_subset <- mnf_mdf[51:64, ]

# Create the plot
mnf<-ggplot(mnf_mdf_subset, aes(x = time)) + 
  geom_line(aes(y = Channel3_Mean, color = "Mean")) + 
  geom_line(aes(y = Channel3_Median, color = "Median")) +
  scale_color_manual(values = c("Mean" = "blue", "Median" = "red"), name=NULL) +
  labs(title = "",  
       x = "Time (s)",
       y = "Value",
       color = "Legend") +
  theme_minimal()+
theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white"))


ggsave("../plot/mnf_3_5.png", plot = mnf, dpi = 300, width = 8, height = 6)


```
#Plot mnf, and mdf for channel 4 for Grip gesture
```{r}
library(ggplot2)

# Filter the first 60 rows for plotting
mnf_mdf_subset <- mnf_mdf[51:64, ]

# Create the plot
mnf<-ggplot(mnf_mdf_subset, aes(x = time)) + 
  geom_line(aes(y = Channel4_Mean, color = "Mean")) + 
  geom_line(aes(y = Channel4_Median, color = "Median")) +
  scale_color_manual(values = c("Mean" = "blue", "Median" = "red"), name=NULL) +
  labs(title = "",  
       x = "Time (s)",
       y = "Value",
       color = "Legend") +
  theme_minimal()+
theme(plot.title = element_text(hjust = 0.5), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.major = element_line(color = "lightgrey"), 
        panel.grid.minor = element_line(color = "white"))


ggsave("../plot/mnf_4_5.png", plot = mnf, dpi = 300, width = 8, height = 6)

```












